---
title: "Assignment:1 "
subtitle: "732A97 Multivariate Statistical Methods,Linkoping University"
author: |
  | **Group 18**
  | Dimitra Muni - dimmu472
  | Karthikeyan Devarajan  - karde799
  | Gowtham KM - gowku593
  | Biswas Kumar - bisku859
#date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggcorrplot)
library(RMaCzek)
```

## Question 1: Describing individual variables  
### a) *Describe the 7 variables with mean values, standard deviations e.t.c.*  
The mean value are  
```{r ,echo=FALSE}
sf <- read.csv('./T1-9.csv')
sf1 <- sf[,2:8]
mean = colMeans(sf[sapply(sf, is.numeric)])
mean
variance = var(sf[,-1])
```
  The standard Deviations are
```{r,echo=FALSE}
deviation = apply(as.matrix(sf[,-1]),2, sd)
deviation
```
  
### b) *Illustrate the variables with different graphs (explore what plotting possibilities R has). Make sure that the graphs look attractive (it is absolutely necessary to look at the labels, font sizes, point types). Are there any apparent extreme values? Do the variables seem normally distributed? Plot the best fitting (match the mean and standard deviation, i.e. method of moments) Gaussian density curve on the data's histogram. For the last part you may be interested in the hist() and density() functions*  
```{r ,echo=FALSE,out.height='15%'}
 x <- colnames(sf[,2:8])
 for(i in 2:8){
  h <-hist(sf[,i], breaks=10, col="red", xlab=x[i],
           main="Histogram")
  xfit <- seq(min(sf[,i]),max(sf[,i]),length=15)
  yfit <- dnorm(xfit,mean=mean(sf[,i]),sd=sd(sf[,i]))
  yfit <- yfit*diff(h$mids[1:2])*length(sf[,i])
  lines(xfit, yfit, col="blue", lwd=2)
}
```
  There are more extreme values in marathon histogram graph. All are normally distributed and with increase in race track distance the gaussian distribution fit curve are skewed towards negative.  The  deviation is more for marathon then 100m race. When a model is created, the accuracy will good for less skewed model. so, the accuracy will be better for 100m racetrack. 
  
## Question 2: Relantionship with the variables  
### a) *Compute the covariance and correlation matrices for the 7 variables. Is there any apparent structure in them? Save these matrices for future use*  

```{r ,echo=FALSE}
sf_variance <- var(sf[,2:8])
sf_variance
sf_correlation <- cor(sf[,2:8])
sf_correlation
  ggcorrplot::ggcorrplot(sf_correlation, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle",
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlation between different RaceTrack")

```
  
  The correlation is decreasing for the racetrack with increase in distance of the racetrack. There is graph to represent the correlation matrix.    

### b) Generate and study the scatterplots between each pair of variables. Any extreme values?   
  
```{r scatterplot,out.height='15%',echo=FALSE}
colna <- colnames(sf)
for(i in 2:8)
   for(j in 2:8)
       if(j!=i & j>i){
       plot(x = sf[,i], y = sf[,j],col = "red", pch = 19,cex = 1,lty = "solid",lwd = 2,xlab = colna[i],ylab = colna[j]) 
       text(x =  sf[,i],y = sf[,j],labels = as.character(sf[,1]), pos = 1)
       }
```
  
  In most of the graphs between the variables, COK,SAM and PNG are the countries appears to extreme countries. 

### c) *Explore what other plotting possibilities R offers for multivariate data. Present other (at least two) graphs that you find interesting with respect to this data set.*  

  
```{r ,echo=FALSE}
sf$X100m_z <- round((sf$X100m - mean(sf$X100m))/sd(sf$X100m), 2)  
sf$X100m_type <- ifelse(sf$X100m_z < 0, "below", "above") 
sf <- sf[order(sf$X100m_z), ]
sf$countries <- factor(sf$countries, levels = sf$countries)
ggplot(sf, aes(x=countries, y=X100m_z, label=X100m_z)) + 
  geom_bar(stat='identity', aes(fill=X100m_type),width = 0.75)  +
  scale_fill_manual(name="Speed", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(subtitle="Normalised Speed of 100 meters of Different Countries", 
       title= "Diverging Bars") + xlab("countries") + ylab("100m") + 
  coord_flip() +  theme(axis.text.x = element_text(size=10),
                        axis.text.y = element_text(size=5))

```

```{r echo=FALSE}
ggplot(sf, aes(x=sf$countries, y=sf$X100m)) + 
  geom_point(col="tomato2", size=3) +   
  geom_segment(aes(x=sf$countries, 
                   xend=sf$countries, 
                   y=min(sf$X100m), 
                   yend=max(sf$X100m)), 
               linetype="dashed", 
               size=0.1) +  
  labs(title="Dot Plot", 
       subtitle="X100m vs Countries") +  xlab("countries") + ylab("100m") +
  coord_flip() + theme(axis.text.x = element_text(size=10),
                    axis.text.y = element_text(size=5))
```


## Question 3: Examining for extreme values

### a) *Look at the plots (esp.scatterplots) generated in the previous question. Which 3-4 countries appear most extreme? Why do you consider them extreme?*

In previous plot of 100m,the countries peformed relatively better were Cook's Island(COK) and Samoa(SAM) while countries performing relatively poor were USA,France,Russia

Extreme countries lies between  2$\sigma$ and 3$\sigma$ away from the $\mu$ in each direction.


### b) *Finding five most extreme countries using Euclidean Distance*


```{r,echo=FALSE,eval=TRUE}
centered=scale(x = as.matrix(sf[,2:8]), center = TRUE, scale = FALSE)
squared=centered%*%t(centered)
rownames(squared)<-sf[,1]
colnames(squared)<-sf[,1]
diagonal<-diag(squared)
extrems=order(diagonal, decreasing=TRUE)[1:5]
cat("Countries which have the extreme values",as.character(sf[extrems,1]))
cat("Sweden's rank",which(names(diagonal[order(diagonal, decreasing=TRUE)])=="SWE"))
```


### *c) Finding five most extreme countries using Normalized Euclidean Distance*

```{r,echo=FALSE,eval=TRUE}
centered=scale(x = as.matrix(sf[,2:8]), center = TRUE, scale = FALSE)
variance<-var(sf[,2:8])
V<-variance*diag(7)
distance2<-centered%*%solve(V)%*%t(centered)
diagonal<-diag(distance2)
names(diagonal)<-sf[,2:8]
extrems=order(diagonal, decreasing=TRUE)[1:5]
cat("Countries which have the extreme values",as.character(sf[extrems,1]))
cat("Sweden's rank",which(names(diagonal[order(diagonal, decreasing=TRUE)])=="SWE"))
```


### *d) Finding five most extreme countries using Mahalanobis distance*
```{r,echo=FALSE,eval=TRUE}
centered=scale(x = as.matrix(sf[,2:8]), center = TRUE, scale = FALSE)
covariance_mat=cov(sf[,2:8])
mahalnob<-centered%*%solve(covariance_mat)%*%t(centered)
diagonal<-diag(mahalnob)
names(diagonal)<-sf[,1]
extrems=order(diagonal, decreasing=TRUE)[1:5]
cat("Countries which have the extreme values",as.character(sf[extrems,1]))
cat("Sweden's rank",which(names(diagonal[order(diagonal, decreasing=TRUE)])=="SWE"))
```




### e) *Compare the results in b) and d),Some of the countries are in the upper end with all the measures and perhaps they can be classified as extreme.How does Sweden behave?*. 
   
Countries performing extremly well or extremly poorly can be classified as extreme from the data.Samoa(SAM),Cook's Island(COK),Papua New Guinea(PNG) have relatively poor performance in the Olympics.
Countries such as Russia and USA have performed well, still all these countries are extreme in terms of their distance from the mean.

Sweden is ranking 48,50 and 54 accordingly for **(b),(c) and (d)**  which suggests that for Sweden the dispersion is less, the data is spread evenly from mean.Sweden has relatively consistent records in all different forms of competitions.
   
   
*Produce Czekanowski's diagram using e.g. the RMaCzek package. *
   
```{r,echo=FALSE,eval=TRUE,out.height='500%',out.width='110%',fig.align='center'}
suppressWarnings(suppressMessages(library(RMaCzek)))
#Czekanowski's diagram
rownames(sf)<-sf[,1]
x<-as.matrix(sf[,2:8])
mat<-czek_matrix(x)
plot(mat)
```
*In case of problems please describe them*

The input dataframe had to be conditioned in a such a way that for the data, rownames were column1.For the function **czek_matrix()** in input data.frame first column was excluded.


# Code Appendix
```{r ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```

