---
title: "Assignment 3"
author: "Karthikeyan Devarajan"
date: "02/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FactoMineR)
library(psych)
sf <- read.csv(file.choose(),header = F)
colnames(sf) <- c("Countries","X100M","X200M","X400M",
                    "X800M","X1500M","X3000M","Marathon")
sf1 <- sf[,-1]
```
  
# Question 1: Principal components, including interpretation of them.  

*The data on national track records for women are listed in Table 1.9.*  
*(a) Obtain the sample correlation matrix R for these data, and determine its eigenvalues and eigenvectors*  

```{r,echo=FALSE}
c_names <- sf$Countries
Column_means <- colMeans(sf1)
vari <- var(sf1)
corr_matrix <- cor(sf1)
eigen_variable <-eigen(corr_matrix)
cat("The correlation Matrix of the data is \n")
corr_matrix
cat("The eigen values of the data is \n",eigen_variable$values,"\n")
#eigen_variable$values
cat("The eigen vector of the data is \n")
eigen_variable$vectors

```
  
  *(b) Determine the first two principal components for the standardized variables. Prepare a table showing the correlations of the standardized variables with the components, and the cumulative percentage of the total (standardized) sample variance explained by the two components.*  

```{r, echo=FALSE}
pca_model <- PCA(sf1,scale.unit = T,ncp = 7,graph = F)
cat("The Percentage of sample variance of PC 1:",pca_model$eig[1,2],"\n")
cat("The Percentage of sample variance of PC 2:",pca_model$eig[2,2],"\n")
cat("The Cumulative Percentage of sample variance of 2 PCA:",pca_model$eig[2,3],"\n")
sf$pcscore1 <- pca_model$ind$coord[,1]
sf$pcscore12 <- pca_model$ind$coord[,2]
sf <- sf[order(sf$pcscore1,decreasing = T),]
cat("The Rank of nation based on PC1:\n")
RA <- sf[,c("Countries","pcscore1")]

```

```{r}
MLfactanal <- factanal(sf1,factors = 2,scores = "regression",control = "lower")
plot(MLfactanal$scores, pch="",main="Factor scores for ML factor analysis")
text(MLfactanal$scores, labels = as.character(sf[,1]),cex=0.7)

```

```{r}
MLfact2 <- factanal(sf1,covmat = NULL,factors  = 3)
load <- MLfact2$loadings[,1:2]
plot(load,type="n") 
text(load,labels=as.character(sf[,1]),cex=.7)

MLfact3 <- factanal(cor(sf1),covmat = var(sf1),n.obs=54,factors  = 3)
MLfact3
root_covar1 <- principal(var(sf[,2:8]), nfactors = 3, rotate = 'varimax', covar = FALSE)
root_covar1
root_covar2 <- principal(cor(sf[,2:8]), nfactors = 3, rotate = 'varimax', covar = FALSE)
root_covar2
```
